---
title: "Reproducible Document"
author: "rostools" 
format: html
---

```{r setup}
#| message: false
#| warning: false

library(tidyverse)

post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, Group, auc_pg, auc_ins, glykemi)
```

```{r}
2 + 2
```
## About me 

- Katrine Bovién Gørlitz 
- Medicinsk Diagnostic Center 
- Central Denmark Hosiptal Region 

I am doing a *PhD* in **Diabetes Technology** 

## Simple code 
```{r}
3 * 3
```

## Testing for Git

## Showing the data 
```{r}
post_meal_data

glimpse(post_meal_data)
```

## Plot one continuous variable 

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram(bins = 20)
```

See @fig-bmi-histo


## Exercise: discrete plots

```{r}
#| fig-cap: "Distribution of glykemi"
#| label: fig-glykemi-histo
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

See @fig-glykemi-histo 


## Exercise: Plotting two discrete variables 

```{r}
#| fig-cap: "Distribituon of BMI and number of patients with glycemia"
#| label: fig-bmi-glykemi-longitudinelt
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

See @fig-bmi-glykemi-longitudinelt

## Side by side plots 

```{r fig_bmi_glycemia}
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#|  - "Distribution of BMI"
#|  - "Number of those with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```


## Plotting two continuous variables 

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

See @fig-bmi-auc-pg


##Selecting data 

```{r}
select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```


## Renaming columns 

```{r}
small_post_meal_data |> 

    rename(
      auc_glucose = auc_pg,
      auc_insulin = auc_ins)
```

