---
title: "Reproducible Document"
author: "rostools" 
format: html
---

```{r setup}
#| message: false
#| warning: false

library(tidyverse)

post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |>
  filter(glykemi == 0) |>
  select(-glykemi)

tidied_post_meal_data <- tidier_post_meal_data |>
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Control", "First-degree relative")
  )
```

```{r}
2 + 2
```
## About me 

- Katrine Bovién Gørlitz 
- Medicinsk Diagnostic Center 
- Central Denmark Hosiptal Region 

I am doing a *PhD* in **Diabetes Technology** 

## Simple code 
```{r}
3 * 3
```

## Testing for Git

## Showing the data 
```{r}
post_meal_data

glimpse(post_meal_data)
```

## Plot one continuous variable 

```{r}
#| fig-cap: "Distribution of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram(bins = 20)
```

See @fig-bmi-histo


## Exercise: discrete plots

```{r}
#| fig-cap: "Distribution of glykemi"
#| label: fig-glykemi-histo
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

See @fig-glykemi-histo 


## Exercise: Plotting two discrete variables 

```{r}
#| fig-cap: "Distribituon of BMI and number of patients with glycemia"
#| label: fig-bmi-glykemi-longitudinelt
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

See @fig-bmi-glykemi-longitudinelt

## Side by side plots 

```{r fig_bmi_glycemia}
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#|  - "Distribution of BMI"
#|  - "Number of those with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```


## Plotting two continuous variables 

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose"
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

See @fig-bmi-auc-pg


##Selecting data 

```{r}
select(post_meal_data, Age, BMI, Weight)

select(post_meal_data, -Group)

select(post_meal_data, starts_with("PG"))

select(post_meal_data, ends_with("DXA"))

select(post_meal_data, contains("0"))
```


## Renaming columns 

```{r}
small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```


## Select and rename tasks

```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("..."))

# Task 3.
post_meal_data |>
  rename(
    P.Glucose.minus.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )

# Task 4.
post_meal_data |>
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)
```


## Filtering data 

```{r}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR")

small_post_meal_data |>
  filter(Group == "CTR", BMI >= 25) # komma (,) virker i dette tilfælde som et &. Bar (|) virker som eller
```

##Modify data 

```{r}
tidier_post_meal_data |>
  mutate(insulin_glucose_ratio = auc_insulin / auc_glucose)
```

## Practice with filter and mutate

# Task 1.
```{r}
post_meal_data |>
  filter(BMI >= 25 & BMI <= 40 & Group == "FDR")
```
#Svar tast 1 = 8 deltagere 


# Task 2.
```{r}
post_meal_data |>
  mutate(
    percent_body_fat = (Fat.mass.DXA / Weight) * 100
  ) |>
  filter(percent_body_fat > 30)
```
#Svar task 2 = 10 deltagere 


## Visualizing with a jitter plot 

```{r}
tidied_post_meal_data |>
  mutate(overweight = if_else(BMI >= 25, "Overweight", "Not overweight")) |>
  ggplot(aes(x = overweight, y = insulin_glucose_ratio)) +
  geom_jitter()
```

## Plotting three or more variables 

```{r}
tidied_post_meal_data |>
  filter(BMI < 30) |>
  mutate(young = if_else(Age < 40, "Young", "Old")) |>
  ggplot(aes(x = BMI, y = insulin_glucose_ratio, colour = Group)) +
  geom_point() +
  facet_grid(cols = vars(young)) +
  labs(
    x = "Body mass index (BMI)",
    y = "Insulin to glucose rato"
  )
```
